---
title: "Employee Involvement"
subtitle: ''
author: Gianna LaFrance, Jaiden Neff, & Samantha Gouveia
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_mono_light(
  header_font_google = google_font("Open Sans", "300"),
  text_font_google   = google_font("Lato", "400"),
  code_font_google   = google_font("IBM Plex Mono"),
  base_color = "#1F7799")

```


<h1 align = "center"> Table of Contents </h1>
<center><font size="7">  
- Introduction <br>
- Exploratory Data Analysis <br>
- Principal Componenet Analysis <br>
- Multinomial Logistic Regression <br>
- Conclusion & Discussion
</font></center>

---

## Introduction

Our goal for this study is to see the effect that factors have on the overall job involvement at our company. We plan to run statistical tests and observe the data to draw conclusions that will lead us to make informed decisions about how to improve job involvement. 

The data set has 1470 rows and 35 columns.  

We will be focusing on the following variables for our exploratory data analysis and we will keep these for our sample population.

- Job Involvement: Likert scale 1-4 from low to very high
- Environment Satisfaction: Likert scale 1-4 from low to very high
- Job Satisfaction: Likert scale 1-4 from low to very high
- Work Life Balance:Likert scale 1-4 from low to very high
- Monthly Income:
- Years Since Last Promotion:
- Years At Company:
- Age:
- Number Companies Worked:
- Department:
- Job Level:
- Years With Current Manager:
- Hourly Rate:


```{r}

employee <- read.csv("https://raw.githubusercontent.com/anafrance/FinalProject/main/EmployeeAttritionData.csv", header=TRUE)

library(tidyverse)
employee1 <- employee %>% 
  dplyr::select(JobInvolvement, JobSatisfaction, EnvironmentSatisfaction, HourlyRate)


```

---
class: inverse center middle

# Exploratory Data Analysis


---

## Job Involvement Data

```{r}

# Print title for frequency distribution of job involvement
cat("Frequency Distribution of Job Involvement:\n")

# Display frequency distribution table of job involvement
print(table(employee$JobInvolvement))


# Frequency graph for Job Involvement
barplot(table(employee$JobInvolvement), main = "Job Involvement", xlab = "Level of Involvement", ylab = "Frequency")

```

---

## Frequency Graphics

```{r}
# Frequency graph for Job Involvement
barplot(table(employee$EnvironmentSatisfaction), main = "Environment Satisfaction", xlab = "Level of Involvement", ylab = "Frequency")
```

---

## Frequency Graphics

```{r}

# Frequency graph for Job Involvement
barplot(table(employee$JobSatisfaction), main = "Job Satisfaction", xlab = "Level of Involvement", ylab = "Frequency")


```


---

## Summary Statistics

```{r}

# Summary statistics with titles for each factor

# HourlyRate
cat("\nSummary statistics for Hourly Rate:\n")
summary(employee$HourlyRate)

# Years Since Last Promotion
cat("\nSummary statistics for Years Since Last Promotion:\n")
summary(employee$YearsSinceLastPromotion)

# Age
cat("\nSummary statistics for Age:\n")
summary(employee$Age)

# Number of Companies Worked For
cat("\nSummary statistics for Number of Companies Worked For:\n")
summary(employee$NumCompaniesWorked)

```

---

## Scatter Plot for Age 

```{r}

# Create a scatter plot of age and job involvement
plot(employee$Age, employee$JobInvolvement,
     xlab = "Age", ylab = "Job Involvement",
     main = "Scatter Plot of Age and Job Involvement")


```

---

## Job Involvemnet and Satisfaction 

```{r}

# Create a contingency table of job involvement and job satisfaction
contingency_table <- table(employee$JobInvolvement, employee$JobSatisfaction)

# Convert the contingency table into a data frame
df <- as.data.frame.matrix(contingency_table)

# Set the row names as the levels of job involvement
row.names(df) <- levels(employee$JobInvolvement)

# Set the colors for the bars
bar_colors <- c("skyblue", "pink", "orange", "green")

# Plot a clustered bar chart
barplot(as.matrix(df), beside = TRUE, col = bar_colors,
        main = "Clustered Bar Chart of Job Involvement vs. Job Satisfaction",
        xlab = "Job Involvement", ylab = "Frequency",
        legend.text = TRUE)

```


---

## Department


```{r}

# Create a barplot to visualize job involvement across departments
barplot(table(employee$Department, employee$JobInvolvement),
        beside = TRUE,
        legend.text = TRUE,
        main = "Job Involvement Across Departments",
        xlab = "Involvemnet",
        ylab = "Frequency",
        col = c("skyblue", "pink", "magenta") # Color of the bars
)

```

---

## Correlation Plot 

```{r}

# Load required libraries
library(corrplot)

# Select relevant columns for correlation analysis
selected_cols <- c("JobInvolvement", "EnvironmentSatisfaction", "JobSatisfaction", 
                   "MonthlyIncome", "YearsSinceLastPromotion", "YearsAtCompany", 
                   "WorkLifeBalance", "Age", "NumCompaniesWorked", "JobLevel", "YearsWithCurrManager", "HourlyRate")

# Create correlation matrix
correlation_matrix <- cor(employee[, selected_cols])

# Plot correlation matrix with clearer labels and larger font size
corrplot(correlation_matrix, method = "circle", type = "upper",
         tl.col = "black", tl.srt = 45,
         title = "Correlation Matrix of Employee Factors",
         tl.cex = 0.8, # Set text label size
         cl.cex = 0.8, # Set color legend label size
         mar = c(1,1,1,1) # Set margin size
)

```

---

## Correlation Coefficents

```{r}

# Calculate correlations between JobInvolvement and other variables
correlation_with_job_involvement <- cor(employee[, selected_cols])["JobInvolvement", ]

# Print correlation coefficients and their significance levels
print(correlation_with_job_involvement)

```


---

## Findings 

We found that there are not that many factors that are highly correlated with Job Involvement there are a few however that have higher correlations and that we will be including in the Final Model 

Multicollinearity in monthly income and job level and Years at Company and Years with Current Manager so we removed these factors because they were not significant on their own 

Monthly Income and Hourly Rate are not correlated which I thought was interesting 

A lot of the variables are more highly correlated to each other than they are to the variable Job Involvement 


---
class: inverse center middle

# Principal Component Analysis


---
name: colors

## PCA

```{r}
eng <- employee
library(tidyverse)
engage <- eng %>% 
  dplyr::select(YearsAtCompany,YearsSinceLastPromotion, JobInvolvement, JobSatisfaction, WorkLifeBalance, Age, NumCompaniesWorked,  MonthlyIncome)
```

```{r}
my.mode = function(dataset){
  freq.tbl = table(dataset)
  max.freq.id=which(freq.tbl==max(freq.tbl))
  mode=names(freq.tbl[max.freq.id])
  as.numeric(mode)
}
```

```{r}

My.plotnScree = function(mat, legend = TRUE, method ="factors", main){

    ev <- eigen(cor(mat))    # get eigenvalues
    ap <- parallel(subject=nrow(mat),var=ncol(mat), rep=5000,cent=.05)
    nScree = nScree(x=ev$values, aparallel=ap$eigen$qevpea, model=method)  
    ##
    if (!inherits(nScree, "nScree")) 
        stop("Method is only for nScree objects")
    if (nScree$Model == "components") 
        nkaiser = "Eigenvalues > mean: n = "
    if (nScree$Model == "factors") 
      nkaiser = "Eigenvalues > zero: n = "
    # axis labels
    xlab = nScree$Model
    ylab = "Eigenvalues"
    ##
    par(col = 1, pch = 18, bg=NA)
    par(mfrow = c(1, 1))
    eig <- nScree$Analysis$Eigenvalues
    k <- 1:length(eig)
    plot(1:length(eig), eig, type="b", main = main, 
        xlab = xlab, ylab = ylab, ylim=c(0, 1.2*max(eig)))
    #
    nk <- length(eig)
    noc <- nScree$Components$noc
    vp.p <- lm(eig[c(noc + 1, nk)] ~ k[c(noc + 1, nk)])
    x <- sum(c(1, 1) * coef(vp.p))
    y <- sum(c(1, nk) * coef(vp.p))
    par(col = 10)
    lines(k[c(1, nk)], c(x, y))
    par(col = 11, pch = 20)
    lines(1:nk, nScree$Analysis$Par.Analysis, type = "b")
    if (legend == TRUE) {
        leg.txt <- c(paste(nkaiser, nScree$Components$nkaiser), 
                   c(paste("Parallel Analysis: n = ", nScree$Components$nparallel)), 
                   c(paste("Optimal Coordinates: n = ", nScree$Components$noc)), 
                   c(paste("Acceleration Factor: n = ", nScree$Components$naf))
                   )
        legend("topright", legend = leg.txt, pch = c(18, 20, NA, NA), 
                           text.col = c(1, 3, 2, 4), 
                           col = c(1, 3, 2, 4), bty="n", cex=0.7)
    }
    naf <- nScree$Components$naf
    text(x = noc, y = eig[noc], label = " (OC)", cex = 0.7, 
        adj = c(0, 0), col = 2)
    text(x = naf + 1, y = eig[naf + 1], label = " (AF)", 
        cex = 0.7, adj = c(0, 0), col = 4)
}
```

```{r factor loadings}

My.loadings.var <- function(mat, nfct, method="fa"){
   # mat =  data matrix
   # nfct = number of factors or components
   # method = c("fa", "pca"), default = is "fa".
    if(method == "fa"){ 
     f1 <- factanal(mat, factors = nfct,  rotation = "varimax")
     x <- loadings(f1)
     vx <- colSums(x^2)
     varSS = rbind('SS loadings' = vx,
            'Proportion Var' = vx/nrow(x),
           'Cumulative Var' = cumsum(vx/nrow(x)))
     weight = f1$loadings[] 
   } else if (method == "pca"){
     pca <- prcomp(mat, center = TRUE, scale = TRUE)
     varSS = summary(pca)$importance[,1:nfct]
     weight = pca$rotation[,1:nfct]
  }
    list(Loadings = weight, Prop.Var = varSS)
}

```


```{r}
library(GPArotation)
library(psych)
library(nFactors)
library(rmarkdown)
library(knitr)
library(parameters)
library(corrplot)
library(ggcorrplot)
library(ggfortify)

JobInvolvement <- engage
My.plotnScree(mat=JobInvolvement, legend = TRUE, method ="components", 
              main="Determination of Job Involvement")
```

---
name: colors

## Factor Loading

```{r}

Loadings = My.loadings.var(mat=JobInvolvement, nfct=2, method="pca")$Loadings
#
# pca loadings
kable(round(Loadings,3),
  caption="Factor loadings of the first few PCAs and the cumulative proportion
of variation explained by the corresponding PCAs in the employee survey.")
```

---

## Proportion of Variation Explained

```{r}
VarProp = My.loadings.var(mat=JobInvolvement, nfct=2, method="pca")$Prop.Var
# pca loadings
kable(round(VarProp,3),
    caption="Cumulative and proportion of variances explained by each 
    the principal component in the employee survey.")
```

---
class: inverse center middle

# Multinomial Logistic Regression

---
name: colors

## MLR

We used the 4th level as the baseline for Job Involvement.

First, we did a model with all of our variables: 
- Environment Satisfaction: Likert scale 1-4 from low to very high
- Job Satisfaction: Likert scale 1-4 from low to very high
- Work Life Balance:Likert scale 1-4 from low to very high
- Years Since Last Promotion:
- Age:
- Number Companies Worked:
- Department:
- Hourly Rate:

The only variables that were statistically significant were Job Satisfaction, Environment Satisfaction, and Hourly Rate.

---
name: colors

## Model

```{css, echo=FALSE}
pre {
  max-height: 500px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r}
# Fit a multinomial logistic regression model 
library(VGAM)
fit <- vglm(JobInvolvement ~ ., 
            data = employee1, 
            family = multinomial) 

```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 200px;
  overflow-y: auto;
  background-color: inherit;
}
```

```{r, class.output="scroll-100"}
print(summary(fit))
```

---

## Model Assumptions

One assumption that needs to be tested is collinearity betweeen the explanatory varibles. 

```{r}
library(corrplot)

cor <- employee1 %>% 
  dplyr::select(JobInvolvement, JobSatisfaction, EnvironmentSatisfaction, HourlyRate)

corrplot(cor(cor), method = "number",tl.cex = 0.6, bg=)

```


---
name: colors

## Conclusion

Not as many variables were correlated/significant in predicting Job Involvement. 

Some of our variables were even negatively correlated.

Job Satisfaction, Environment Satisfaction, and Hourly Rate were significant at the 0.1 alpha level.

Job Involvement is self reported so we might not have accurate data which would lead to these low correlations because of the bias

Continued work: restructuring PCs to include in the model to get better estimates. 

Suggestions: We could measure Job Involvement in other ways such as hours worked or projects completed in the last 6 months. We could also make the survey anonymous so that people feel more comfortable reporting accurately without fear of losing their job. 

---

## Contributions

Introduction and Exploratory Analysis ~ Jaiden

PCA ~ Samantha

Multinomial Logistic Regression ~ Gianna




