---
title: '<font size = 7 color = "Purple">Multiple Linear Regression</font>'
subtitle:
author: '<font size = 5 color = "Purple"> Jaiden Neff & Josh Grant </font>'
institute: '<font size = 6 color = "Purple">West Chester University of Pennsylvania</font><br> '

output:
  xaringan::moon_reader:
    css: 
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("leaflet")) {
   install.packages("leaflet")
   library(leaflet)
}
if (!require("EnvStats")) {
   install.packages("EnvStats")
   library(EnvStats)
}
if (!require("MASS")) {
   install.packages("MASS")
   library(MASS)
}
if (!require("phytools")) {
   install.packages("phytools")
   library(phytools)
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("nleqslv")) {
   install.packages("nleqslv")
   library(nleqslv)
}
if (!require("pander")) {
   install.packages("pander")
   library(pander)
}
if (!require("psych")) {   
  install.packages("psych")
   library(psych)
}
if (!require("car")) {
   install.packages("car")
   library(car)
}
if (!require("lubridate")) {
   install.packages("lubridate")
   library(lubridate)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("simputation")) {
   install.packages("simputation")
   library(simputation)
}
if (!require("modelr")) {
   install.packages("modelr")
   library(modelr)
}
if (!require("MASS")) {
   install.packages("MASS")
   library(mass)
}
if (!require("pROC")) {
   install.packages("pROC")
   library(pROC)
}
if (!require("readxl")) {
   install.packages("readxl")
   library(readxl)
}
if (!require("forecast")) {
   install.packages("forecast")
   library(readxl)
}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "grey",
  inverse_header_color = "#FFFFFF"
)
```


class:inverse4, top

<h2 align="center"> Table of Contents</h2>
<BR>

.pull-left[
- Data Description 
- Exploratory Analysis
   - Scatter plots of height and foot length 
   - Box plot of Gender 
- Full Model
-Residual plots 
- Transformations
   - Square Root
   - Log Transformation
- Regression Analysis 
- Goodness of Fit
- Boot Strapping
- Final Model 
- Conclusion 
]



---


<h2 align="center"> Data Description </h2>


The primary question is to identify the association between the hand length (y) and the variables that are closely related. 


`*   IdGen`<br>
`*   gender(X1): male or female`<br>
`*   height(X2): height in cm	`<br>
`*   footLen(X3): foot length in cm	`<br>
`*   handLen(Y): hand length in cm`


---

<h2 align="center"> Gender vs Hand length </h2>


```{r}

data <- read.csv("https://raw.githubusercontent.com/jaidenneff/sta490/main/stature_hand_foot%20(1)%20(1).csv", header = TRUE)
#head(diabites)
#dim()

```

```{r echo=FALSE, fig.align='center', out.width = '600px'}

data$gender<-factor(data$gender,
                          levels=c(1,2),
                          labels=c("Male", "Female"))

Genderplot = ggplot(data, aes(x = gender, y = handLen)) + geom_boxplot() 
Genderplot



```


---

<h2 align="center"> Height vs Hand Length  </h2>



```{r echo=FALSE, fig.align='center', out.width = '600px'}

attach(data)
plot(height,handLen,  
 xlab="height ", ylab="handLen ", pch=19)
```

---

<h2 align="center"> Foot Length vs Hand Length</h2>


```{r echo=FALSE, fig.align='center', out.width = '600px'}
attach(data)
plot(footLen,handLen, 
 xlab="footLen ", ylab="handLen ", pch=19)
```

---

<h2 align="center"> Full Model </h2>


```{r echo=FALSE, fig.align='center', out.width = '600px'}
full.model = lm(handLen ~ ., data = data)
kable(summary(full.model)$coef, caption ="Statistics of Regression Coefficients")
```

---

<h2 align="center"> Residual Plot </h2>


```{r fig.align='center', fig.width=6, fig.height=6, fig.cap="Residual plots of the full model", out.width = '600px'}
par(mfrow=c(2,2))
plot(full.model)

```



---

<h2 align="center"> Square root transformation </h2>


```{r}
sqrt.length.log.dist = lm((handLen)^0.5 ~ height + footLen + gender,  data = data)
kable(summary(sqrt.length.log.dist)$coef)
```



<h2 align="center"> Log transformation </h2>

```{r}
log.length = lm(log(handLen) ~ height + footLen + gender, data = data)
kable(summary(log.length)$coef)
```




---
#Regression Analysis
```{r ,fig.align='center', fig.width= 7, fig.height=4, echo=FALSE}
#define plotting area
par(pty = "s", mfrow = c(1, 3))
#Q-Q plot for original model
qqnorm(full.model$residuals, main = "Full-Model")
qqline(full.model$residuals)
#Q-Q plot for Box-Cox transformed model
qqnorm(log.length$residuals, main = "Log-length")
qqline(log.length$residuals)
#display both Q-Q plots
qqnorm(sqrt.length.log.dist$residuals, main = "sqrt length log dist")
qqline(sqrt.length.log.dist$residuals)
```



---

# Goodeness Of Fit

```{r}
select=function(m){ # m is an object: model
 e = m$resid                           # residuals
 n0 = length(e)                        # sample size
 SSE=(m$df)*(summary(m)$sigma)^2       # sum of squared error
 R.sq=summary(m)$r.squared             # Coefficient of determination: R square!
 R.adj=summary(m)$adj.r                # Adjusted R square
 MSE=(summary(m)$sigma)^2              # square error
 Cp=(SSE/MSE)-(n0-2*(n0-m$df))         # Mellow's p
 AIC=n0*log(SSE)-n0*log(n0)+2*(n0-m$df)          # Akaike information criterion
 SBC=n0*log(SSE)-n0*log(n0)+(log(n0))*(n0-m$df)  # Schwarz Bayesian Information criterion
 X=model.matrix(m)                     # design matrix of the model
 H=X%*%solve(t(X)%*%X)%*%t(X)          # hat matrix
 d=e/(1-diag(H))                       
 PRESS=t(d)%*%d   # predicted residual error sum of squares (PRESS)- a cross-validation measure
 tbl = as.data.frame(cbind(SSE=SSE, R.sq=R.sq, R.adj = R.adj, Cp = Cp, AIC = AIC, SBC = SBC, PRD = PRESS))
 names(tbl)=c("SSE", "R.sq", "R.adj", "Cp", "AIC", "SBC", "PRESS")
 tbl
 }

```

```{r}
output.sum = rbind(select(full.model), select(sqrt.length.log.dist), select(log.length))
row.names(output.sum) = c("full.model", "sqrt.length.log.dist", "log.length")
kable(output.sum, caption = "Goodness-of-fit Measures of Candidate Models")
```
---
#Bootstrapping regression
```{r fig.width=7, fig.height=5}
log.length = lm(log(handLen) ~ height + footLen + gender, data = data)
##
B = 1000     
## 
num.p = dim(model.frame(log.length))[2]  
smpl.n = dim(model.frame(log.length))[1] 
coef.mtrx = matrix(rep(0, B*num.p), ncol = num.p)       
## 
for (i in 1:B){
  bootc.id = sample(1:smpl.n, smpl.n, replace = TRUE) 
  
  log.length.btc = lm(log(handLen) ~ height + footLen + gender, data = data [bootc.id,])     
  coef.mtrx[i,] = coef(log.length.btc)    # extract coefs from bootstrap regression model    
}

cmtrx <- summary(log.length)$coef
```
```{r}
boot.hist = function(cmtrx, bt.coef.mtrx, var.id, var.nm){
 
  x1.1 <- seq(min(bt.coef.mtrx[,var.id]), max(bt.coef.mtrx[,var.id]), length=300 )
  y1.1 <- dnorm(x1.1, mean(bt.coef.mtrx[,var.id]), sd(bt.coef.mtrx[,var.id]))
  highestbar = max(hist(bt.coef.mtrx[,var.id], plot = FALSE)$density) 
  ylimit <- max(c(y1.1,highestbar))
  hist(bt.coef.mtrx[,var.id], probability = TRUE, main = var.nm, xlab="", 
       col = "azure1",ylim=c(0,ylimit), border="lightseagreen")
  lines(x = x1.1, y = y1.1, col = "red3")
  lines(density(bt.coef.mtrx[,var.id], adjust=2), col="blue") 
 
}
```
```{r fig.align='center', fig.width=7, fig.height=5,out.width = '750px'}
par(mfrow=c(2,2))  # histograms of bootstrap coefs
boot.hist(bt.coef.mtrx=coef.mtrx, var.id=1, var.nm ="Intercept" )
boot.hist(bt.coef.mtrx=coef.mtrx, var.id=2, var.nm ="Height" )
boot.hist(bt.coef.mtrx=coef.mtrx, var.id=3, var.nm ="Footlen" )
boot.hist(bt.coef.mtrx=coef.mtrx, var.id=4, var.nm ="gender" )

```

---
## Bootstrapping Residuals

```{r}
num.p = dim(coef.mtrx)[2]  # number of parameters
btc.ci = NULL
btc.wd = NULL
for (i in 1:num.p){
  lci.025 = round(quantile(coef.mtrx[, i], 0.025, type = 2),8)
  uci.975 = round(quantile(coef.mtrx[, i],0.975, type = 2 ),8)
  btc.wd[i] =  uci.975 - lci.025
  btc.ci[i] = paste("[", round(lci.025,4),", ", round(uci.975,4),"]")
 }
#as.data.frame(btc.ci)
kable(as.data.frame(cbind(formatC(cmtrx,4,format="f"), btc.ci.95=btc.ci)), 
      caption = "Regression Coefficient Matrix")
```

---

## Final Model

```{r}
## Final model

model.resid = log.length$residuals
##
B=1000
num.p = dim(model.matrix(log.length))[2]   # number of parameters
samp.n = dim(model.matrix(log.length))[1]  # sample size
btr.mtrx = matrix(rep(0,4*B), ncol=num.p) # zero matrix to store boot coefs
for (i in 1:B){
  ## Bootstrap response values
  bt.lg.length = log.length$fitted.values + 
        sample(log.length$residuals, samp.n, replace = TRUE)  # bootstrap residuals
  # replace PriceUnitArea with bootstrap log price
  data$bt.lg.length =  bt.lg.length   #  send the boot response to the data
  btr.model = lm(bt.lg.length ~ height + footLen + gender, data = data)
  
  btr.mtrx[i,]=btr.model$coefficients
}
```

$$
log(handLen) = 4.3042058 +(0.0005444* height) + (0.0003162 *footLen) - (0.0057154* genderFemale)
$$
---
# Final Thoughts

• The log transformation Is the best model for determining correlation between hand length 
and closely related variables

• Regression analysis wasn’t helpful 

• Bootstrapping isn’t necessary for the final model because the data is normal

• Kept the hand length and gender even though they weren’t “highly correlated” because 
they are important and you can see in exploratory analysis that it has correlation.
